Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-04 paper/sec-analysis.tex
more work on paper

The basic idea of present analysis is similar to the
scale-invariant resonance tagging introduced
in Ref.~\cite{Gouzevitch:2013qca}.
%
rather than restricting to a specific event topology, such as boosted
or resolved, we aim to consistently combine the information from
the three possible event topologies: boosted, intermediate and
resolved.
%
However, an important difference compared to~\cite{Gouzevitch:2013qca}
is that here each category is optimised separately, and then
a exclusive categorisation is defined starting from the category
with the highest signal significance.
%
An advantage of our approach is that the analysis strategy would be unchanged
if the underlying production model of Higgs pairs is modified, for example in the case of
BSM dynamics, which can substantially increase the degree of boost of the final state.

Let us therefore define the three event categories that
we use:
\begin{itemize}
\item {\it Boosted category}.

  An event is assigned to the {\it boosted} category if it
  contains at least two large-$R$ jets, defined as above,
  and where  the two leading large-$R$ jets have been
   double-$b$-tagged.
  %
 Therefore, each of these two large-$R$ double-$b$-tagged jets corresponds to the
 decay products of one Higgs boson candidate.

@@

  An event is assigned to the {\it intermediate} category if
  one and only one large $R$ double-$b$-tagged jet is present,
  which corresponds to the the first Higgs candidate.
  %
  In addition, we require at least two small-$R$ jets,
  which must be separated with respect to the large-$R$
  jet by at least an angular separation of $\Delta R\ge 1.0$.
  %
  These small-$R$ jets must be $b$-tagged as well.

  The second Higgs boson candidate is reconstructed
  by selecting the two $b$-tagged small-$R$ jets that minimize the difference
  between the invariant mass of the large-$R$ jet,
  with the invariant mass of the dijet obtained
  from the sum of the two small-$R$ jets.

\item {\it Resolved category}.
  %
  An event is assigned to the {\it resolved} category
  if at least
  four $b$-tagged small-$R$ jets are present.
  %
  The two Higgs candidates are then reconstructed out of the
  leading four small-$R$ jets in the event
  by considering all possibilities of forming two pairs of jets
  and then choosing the configuration that minimizes the relative difference of
  dijet masses.
  %
  The requirement of restricting this selection to only the four
  leading small-$R$ jets arises from the need to reducing the
  combinatorics from  fakes.
\end{itemize}
These three categories are in principle not exclusive:
a given even can be assigned to more than one category, for
example, satisfying both the requirements of the boosted and resolved
categories at the same time.
%
The exception is the boosted and intermediate categories, which have
orthogonal jet selection requirements.
%
The separate optimisation of the three categories needs to be performed in
a inclusive analysis.
%
Since in general there will be a certain degree of overlap between categories,
a priorisation needs to be introduced in a exclusive analysis.
%

Once a Higgs boson candidate has been identified,
its invariant mass is required to be within a fixed window
around the nominal Higgs boson mass.
%
In particular we require the condition
\be
\label{higgsmasswindow}
|m_{h,j} - 125| < 40~{\rm GeV} \, ,\, j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of each of the two reconstructed  Higgs candidates.
%
This cut is substantially looser as the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}: the motivation
for this choice is that the MVA will take care of the optimisation of these
and other cuts that are not explicitly performed.
%
Only events where the two Higgs candidates satisfy the
condition in Eq.~(\ref{higgsmasswindow}) are classified as signal events.
%
In Table~\ref{sec:categorisation} we summarize the definitions of each
category.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{c|c|c|c}
    \hline
    &  Boosted  &  Intermediate  &  Resolved \\
    \hline
    \hline 
    Jet selection  &  $\ge 2$ large-$R$ jets  & 1 large-$R$ jet, $\ge 2$ small-$R$
    jets  &  $\ge$ 4 small-$R$ jets \\[0.2cm]
    \hline
    Jet cuts  & $p_T \ge 200$ GeV   &  $p_T \ge 200$ GeV (large-$R$ jet)
    &  $p_T \ge 40$ GeV \\
    &  & $p_T \ge 40$ GeV (small-$R$ jets)   &    \\
    &   $|\eta|<2.0$  & $|\eta|<2.0$ (large-$R$ jet) & $|\eta|<2.5$ \\
    &        & $|\eta|<2.5$ (small-$R$ jet)  &  \\[0.2cm]
    \hline
    $b$-tagging  & 2 double-$b$-tags  & 1 double-$b$-tag & 4 single-$b$-tags \\
    &  & and 2 single-$b$-tags  &  \\
    & and $\ge 2$ small-$R$ subjets  &  and $\ge 2$ small-$R$ subjets  &  \\
     &  in large-$R$ jets  &  in large-$R$ jet &  \\
    &  $p_T^{\rm sj}\ge 50$ GeV & $p_T^{\rm sj}\ge 50$ GeV   &  \\[0.2cm]
    \hline
    Mass-Drop Tag  &  Two leading large-$R$ jets  &  Large-$R$ jet &   \\
    \hline
    \end{tabular}
  \caption{\small Definition of the three event categories used in this
    work, together with the corresponding basic kinematical cuts.
    %
    The categories are not exclusive, and each event can satisfy the
    requirements of more than one category at the same time.  
\label{sec:categorisationTable}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The selection requirements and the kinematic cuts
for each of the categories in Table~\ref{sec:categorisationTable} have
been optimised separately.
%
First of all we have performed a inclusive analysis, and optimised the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with higher significance turns out to be the boosted,
followed by the intermediate and the resolved ones.
%
Therefore for the final analysis we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is kept for
these category, else we check if it suits the intermediate
requirements, and failing also this, the event is attempted to
be classified in the
resolved category.
%
The resulting exclusive event samples are then separately processes
through the MVA.
%
This allows us to combine the significance of the dominant topology
(boosted) with that of the sub-leading categories, intermediate
and resolved, without the problem of the double counting.

>>>>

The basic idea of the present analysis is similar to the
scale-invariant resonance tagging introduced
in Ref.~\cite{Gouzevitch:2013qca}.
%
Rather than restricting to a specific event topology,
we aim to consistently combine the information from
the three possible event topologies: boosted, intermediate and
resolved.
%
The optimal cuts for each category will be determined
separately, and then
a exclusive categorisation is defined starting from the category
with the highest signal significance.
%
An advantage of our approach is that the analysis strategy
is robust under variations of
the underlying production model of Higgs pairs is modified,
for instance in the case of
BSM dynamics, which can substantially increase
the degree of boost of the final state.
%
These three categories are defined as follows:
\begin{itemize}
\item {\it Boosted category}.

  An event which
  contains at least two large-$R$ jets, with the two leading
 being $b$-tagged.
 %
 Each of these two large-$R$ $b$-tagged jets are is this
 a candidate
 to contain the decay products of a Higgs boson.

@@

  An event with only one large-$R$ $b$-tagged jet, which
  is assigned to be the leading Higgs candidate.
  %
  In addition, we require at least two small-$R$ $b$-tagged jets,
  which must be separated with respect to the large-$R$ jet
  by $\Delta R\ge 1.0$.
  %

  The subleading Higgs boson candidate is then reconstructed
  by selecting the two $b$-tagged small-$R$ jets that minimize the difference
  between the invariant mass of the large-$R$ jet
  with the invariant mass of the dijet obtained
  from the sum of the two small-$R$ jets.

\item {\it Resolved category}.

 An event with at least
  four $b$-tagged small-$R$ jets.
  %
  The two Higgs candidates are reconstructed out of the
  leading four small-$R$ jets in the event
  by considering all possibilities of forming two pairs of jets
  and then choosing the configuration that minimizes the relative difference of
  dijet masses.
  %

Once a Higgs boson candidate has been identified,
its invariant mass is required to lie within a fixed window
around the nominal Higgs boson mass.
%
In particular we require the condition
\be
\label{higgsmasswindow}
|m_{h,j} - 125| < 40~{\rm GeV} \, ,\, j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of each of the two reconstructed  Higgs candidates.
%
This cut is substantially looser as the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}: the motivation
for this choice is that the MVA will take care of the optimisation of these
and other cuts that are not explicitly performed.
%
Only events where the two Higgs candidates satisfy
Eq.~(\ref{higgsmasswindow}) are classified as signal events.
%

These three categories are not exclusive:
a given event can be assigned to more than one category, for
example, satisfying the requirements of both the intermediate
and resolved
categories at the same time.
%
The exception is the boosted and intermediate categories, which have
orthogonal jet selection requirements.
%
The separate optimisation of the three categories needs to be performed in
a inclusive analysis, and then we introduce a
 priorisation starting with the category of highest signal significance.

 This is achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with higher significance turns out to be the boosted,
followed by the intermediate and the resolved ones, the two latter
with similar significance.
%
Therefore for the final analysis we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is kept for
these category, else we check if it suits the intermediate
requirements, and failing also this, the event is attempted to
be classified in the
resolved category.
%
The resulting exclusive event samples are then separately processes
through the MVA.
%
This allows a consistent combination
of the significance of the three event categories.
